---
title: "Clean merged data sets and run regression working paper"
author: "Daniel Malone"
date: "2023-11-06"
output: html_document
---


Set up chunk with libraries which may be useful.
```{r}
library(tidyverse)
library(zoo)
library(future)
library(future.apply)
library(tictoc)
library(did)
plan(multisession)
availableCores()
setwd("C:/Users/malon/OneDrive/Documents/GitHub/econometrics-project-B/Econometrics Project/Data")
```

Chunk which cleans data and creates conditions which determine drought, based on yearly precipitation patterns
```{r}
#make a function which will clean the data set, and add relevant variables
cleanmaster = function(x){
#make a copy of the master data set, then create two new colums which record the state fip code and county fip code separately
  master2 = x|>
    mutate(fipscopy = fips)|>
    separate_wider_position(fipscopy, c(statefip = 2, countyfip = 3))

#modifies the year and precipitation variables so that they can be better used for regression analysis
  master3 = master2
  master3$year = factor(as.numeric(master3$year))
  master3$precip = as.numeric(master3$precip)

#create two new variables.  One records last year's precipitation, and the other records whether there was a year to year drop in precipitation
  master3 = master3|>
    mutate(lastyearprecip = lag(precip))|>
    mutate(dropprecip = if_else(precip - lastyearprecip < 0, 1, 0))

#create three new variables, which records if there was a drought in that county that year.  This is recorded as three year droughts, four year droughts, and five year droughts.
  master4 = master3|>
    group_by(fips)|>
    mutate(drought = ifelse(dropprecip == 1 & lag(dropprecip)== 1 & lag(dropprecip, n = 2)==1,1,0))|>
    mutate(droughtlength = ifelse(drought == 1, 1,
                           ifelse(drought == 1 & lag(drought) == 1, 2, 
                           ifelse(drought == 1 & lag(drought) == 1 & lag(drought, n=1) == 1, 3,
                           ifelse(drought == 1 & lag(drought) == 1 & lag(drought, n=2) == 1 & lag(drought, n=3) == 1, 4, 0)))))
#there is a flaw with this method;  it requires at least three years for the data to register a drought, even though a drought may have been occurring during this time period.  It may be useful to pull precipitation data from 4 years prior to the start date, or to simply move the goalposts, and make sure the regression only runs for time periods without false negative errors.
  
  return(master4)
}
```
This chunk runs the specified function above and creates a new dataset which has been cleaned
```{r}
cleandat = cleanmaster(master_data)
```

Create post and pre drought effects(THIS PART IS OUTDATED PLEASE IGNORE)
```{r}
#this creates a variable which records the post drought effects
  cleandat = cleandat|>
    group_by(fips)|>
    mutate(postdrought = ifelse(drought == 0 & lag(drought) == 1, 0,
                                ifelse(drought == 0 & lag(drought, n = 2) == 1, 1,
                                ifelse(drought == 0 & lag(drought, n = 3) == 1, 2,
                                ifelse(drought == 0 & lag(drought, n = 4) == 1, 3,
                                ifelse(drought == 0 & lag(drought, n = 5) == 1, 4, .))))))

  
#this creates a varible which records the pre drought effects
  cleandat = cleandat|>
    group_by(fips)|>
    mutate(predrought = ifelse(drought == 0 & lead(drought, n = 1) == 1, 1,
                                ifelse(drought == 0 & lead(drought, n = 2) == 1, 2,
                                ifelse(drought == 0 & lead(drought, n = 3) == 1, 3,
                                ifelse(drought == 0 & lead(drought, n = 4) == 1, 4,
                                ifelse(drought == 0 & lead(drought, n = 5) == 1, 5, 0))))))
```
 
```{r}  
cleandat = cleandat|>
  group_by(fips)|>
  filter(year != "2000",
         year != "2001")

  cleandat = cleandat|>
    group_by(fips)|>
    mutate(droughtstart = ifelse(drought == 1 & lag(drought) == 0, 1, 0),
           droughttime = 99,
           droughttime = ifelse(droughtstart == 1, 0,
                                ifelse(lag(droughtstart, n = 1) == 1 & is.na(lag(droughtstart, n = 1)) == FALSE , 1,
                                ifelse(lag(droughtstart, n = 2) == 1 & is.na(lag(droughtstart, n = 2)) == FALSE, 2,
                                ifelse(lag(droughtstart, n = 3) == 1 & is.na(lag(droughtstart, n = 3)) == FALSE, 3,
                                ifelse(lag(droughtstart, n = 4) == 1 & is.na(lag(droughtstart, n = 4)) == FALSE, 4,
                                ifelse(lag(droughtstart, n = 5) == 1 & is.na(lag(droughtstart, n = 5)) == FALSE, 5,
                                ifelse(lead(droughtstart, n = 1) == 1 & is.na(lag(droughtstart, n = 1)) == FALSE, -1,
                                ifelse(lead(droughtstart, n = 2) == 1 & is.na(lag(droughtstart, n = 2)) == FALSE, -2,
                                ifelse(lead(droughtstart, n = 3) == 1 & is.na(lag(droughtstart, n = 3)) == FALSE, -3,
                                ifelse(lead(droughtstart, n = 4) == 1 & is.na(lag(droughtstart, n = 4)) == FALSE, -4,
                                ifelse(lead(droughtstart, n = 5) == 1 & is.na(lag(droughtstart, n = 5)) == FALSE, -5, 99 ))))))))))))


#creates a variable which records the migration per population
  cleandat$net_migration = as.numeric(cleandat$net_migration)
  cleandat$pop_estimates = as.numeric(cleandat$pop_estimates)
  cleandat = cleandat|>
    group_by(fips)|>
    mutate(migrationperpop = net_migration / pop_estimates)
  cleandat$predrought = as.numeric(cleandat$predrought)
  cleandat$postdrought = as.numeric(cleandat$postdrought)
```

Checks some things in the dataset, then creates two datasets which are specifically for county level and state level data
```{r}
#counts the number of counties which were in a three year, four year, and five year drought
cleandat|>
  group_by(drought)|>
  count(drought)


#points out which states are in the dataset.
cleandat2 = cleandat
cleandat2$statefip = factor(cleandat2$statefip)
cleandat2|>
  group_by(statefip)|>
  count(statefip)
#3, 7, 14, 43, 52, are skipped.

#determine if states have both precip and net migration data
cleandat2|>
  filter(is.na(precip) == FALSE & is.na(net_migration) == FALSE)|>
  group_by(statefip)|>
  count(statefip)
#3, 7, 14, 43, and 49 lack either precip data or net migration data

#determine if states have precip data
cleandat2|>
  filter(is.na(precip) == FALSE)|>
  group_by(statefip)|>
  count(statefip)
#3, 7, 14, 43, and 49 lack precipitation data

#determine if states have net migration data
cleandat2|>
  filter(is.na(net_migration) == FALSE)|>
  group_by(statefip)|>
  count(statefip)
#3, 7, 14, 43 lack migration data.
```

Creates a new data set, which trims out years outside of the scope of the project, then removes all NA values in relevant fields.
```{r}
#creates two separate datasets off the cleaned data set which separates fips codes that refer to state data and county data.
countymaster = cleandat|>
  filter(countyfip != "000",
         year != "2000",
         year != "2001",
         is.na(precip) != TRUE,
         is.na(net_migration) != TRUE,
         is.na(drought) != TRUE,
         is.na(state) == FALSE,
         is.na(lastyearprecip) != TRUE,
         is.na(year) != TRUE)

#check for droughts and unique fips values
countymaster|>
  group_by(drought)|>
  count(drought)

countymaster|>
  filter(is.na(drought))|>
  count(fips)
```

```{r}
cleandat|>
  group_by(fips)|>
  count(droughtstart)|>
  filter(droughtstart ==1 & n>1)

cleandat|>
  group_by(year)|>
  count(droughtstart,sort = TRUE)
count
```
Save the data to the github folder
```{r}
#set working directory to the github folder.  Remember to change this path to your computer!
wd = ("C:/Users/malon/OneDrive/Documents/GitHub/econometrics-project-B/Econometrics Project/Data")
setwd(wd)

#saves the cleaned master dataset.  commented lines of code are previous versions of the dataset
#write_csv(cleandat,file = "readydat")
#filename = "Rmasterdat.Rda"
#save(cleandat, file = filename)
#filename = "Rmasterdat2.Rda"
#save(countymaster, file = filename)
#write_csv(countymaster, file = "countymaster")
#filename = "Rmasterdat3.Rda"
#save(countymaster, file = filename)
#write_csv(countymaster, file = "countymaster2")
filename = "Rmasterdat4.Rda"
save(countymaster, file = filename)
write_csv(countymaster, file = "countymaster3")
```


```{r}
#runs the regression
# -- WARNING -- 
#It takes a long time to run this regression of code.  When running it, be patient, and let your computer run.  It may also be wise to add code which allows R to run all available cores at some point, if we want to run this line of code for testing purposes.
#this regression takes a significant amount of computational power, that R might not be the optimal use here.  Maybe running the exported readymasterdat file on Stata would work better?

terms = countymaster$fips + countymaster$state * countymaster$year
ols = lm(countymaster$net_migration ~ countymaster$lastyearprecip + countymaster$smalldrought + countymaster$lastyearprecip * countymaster$smalldrought + countymaster$fips + countymaster$state * countymaster$year)
summary(ols)

ols2 = lm(countymaster$net_migration ~ countymaster$lastyearprecip + countymaster$fips + countymaster$year)
summary(ols2)
sc = as.data.frame(ols$coefficients)
sc = rownames_to_column(sc, "names")
sc$names = gsub("countymaster$", "", sc$names)

```